The Weakly Nonlinear Magnetorotational Instability in a local geometry

### 2017-03-14 paper/thingap_ref_response/thingap_mri_paper_ref_response_apj.tex
changed sign of Gtwiddle in thingap text
updated Oishi grant number
added discussion of growth rate with B0; added figure.

%%
%% Beginning of file 'sample61.tex'
%%
%% Modified 2016 September
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.1 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

@@

Because we nondimensionalize $B$ by the magnitude of the background field strength, $B_0 \equiv 1$ in Equations \ref{eqset1} - \ref{eqset4}. To excite the weakly nonlinear MRI, we tune the background magnetic field down away from stability (recall that stronger vertical fields stabilize the MRI). We do so by substituting $B = B_0\left(1 - \epsilon^2\right)$. The degree of departure from the marginal state is measured by the small parameter $\epsilon$. An $\mathcal{O}\left(\epsilon^2\right)$ weakening of the background magnetic field destabilizes a finite band of wave modes with a width of $\mathcal{O}\left(\epsilon\right)$, which interact nonlinearly.

@@

\section{Acknowledgments}
S.E.C. was supported by a National Science Foundation Graduate Research Fellowship under grant No. DGE-16-44869. J.S.O. acknowledges support from NASA grant 15-LWS15\_2-0063. The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.

@@

\beq
\widetilde{\mathcal{G}} = - \mathcal{G} \partial_z - \mathcal{L}_3 \partial_z^3,
\eeq

>>>>

%% Beginning of file 'sample61.tex'
%%
%% Modified 2016 September
%%
%% This is a sample manuscript marked up using the
%% AASTeX v6.1 LaTeX 2e macros.
%%
%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
%% (Copyright 2000-2015).  See the classfile for details.

@@

Because we nondimensionalize $B$ by the magnitude of the background field strength, $B_0 \equiv 1$ in Equations \ref{eqset1} - \ref{eqset4}. To excite the weakly nonlinear MRI, we tune the background magnetic field down away from stability (recall that stronger vertical fields stabilize the MRI). We do so by substituting \edit1{$B = B_0\left(1 + \epsilon^2\right)$}. The degree of departure from the marginal state is measured by the small parameter $\epsilon$. An $\mathcal{O}\left(\epsilon^2\right)$ \edit1{strengthening} of the background magnetic field destabilizes a finite band of wave modes with a width of $\mathcal{O}\left(\epsilon\right)$, which interact nonlinearly. \edit1{We note that this definition of $\epsilon$ is opposite in sign to nearly all previous works \citep[e.g.][]{Umurhan:2007dz,Umurhan:2007hs}. Because in the ideal limit, the MRI can be tuned into instability by setting $B_0$ to its critical value and then \emph{decreasing} its value, it is natural to consider $\epsilon^2$ as a weakening of the background field. However, as we show in figure~\ref{fig:growth_rate_vs_B0}, for the dissipative case with $\eta, \nu \ne 0$, when all other parameters are critical, decreasing $B_0$ leads to stability, while increasing it pushes the system into instability. Figure~\ref{fig:growth_rate_vs_B0} is symmetric about $B_0 = 0$, as it must be, since the MRI is insensitive to the sign of the background field. There are several places at which the derivative of $\gamma$ appears discontinuous; this is not physical but rather reflects the fact that we define $\gamma$ as the growth rate of the most unstable mode. That is, it is the maximum real part of the eigenvalues of the linearized system (e.g. equation~\ref{eq:unperturbed_matrix_equations} with $\mathbf{N} = 0$). Because there are four wave families in rotating incompressible MHD, each modified differently by changing $B_0$, when the growth rates of the individual modes cross, there appear piecewise continuous solutions. We highlight one such point in the inset in figure~\ref{fig:growth_rate_vs_B0}, where the MRI mode becomes more stable than another mode which is always stable. Since all of these piecewise discontinuities are below $\gamma=0$, they do not affect the analysis here.}

\begin{figure}[[h!]]
  \centering
  \includegraphics[width=\columnwidth]{growth_vs_B0.pdf}
  \caption{Growth rate $\gamma$ as a function of background magnetic field strength $B_0$ at $\reym = \reym_c$, $k_z = k_c$, $\Pm = 10^{-3}$. Around the critical value $B_0 = 1.$, \emph{strengthening} $B_0$ tunes the system into instability, while \emph{decreasing} it leads to stability. The inset highlights the fact that $\gamma$ is determined by the maximum real eigenvalue of the system, which switches from one mode family to another as discussed in the text.}
  \label{fig:growth_rate_vs_B0}
\end{figure}

@@

\section{Acknowledgments}
S.E.C. was supported by a National Science Foundation Graduate Research Fellowship under grant No. DGE-16-44869. J.S.O. acknowledges support from NASA grant \edit1{NNX16AC92G}. The authors thank Mordecai Mac Low, Jeremy Goodman, John Krommes, Geoff Vasil, and Ellen Zweibel for useful discussion.

@@

\beq
\widetilde{\mathcal{G}} = \mathcal{G} \partial_z + \mathcal{L}_3 \partial_z^3,
\eeq
